keep the scripts here, to better document what is changing
 load the libraries gapminder and ggplot2 and dplyr and tidyr
```{r}
library(gapminder)
library(ggplot2)
library(dplyr)
library(tidyr)
```

```{r}
ggplot(data = gapminder, aes(x = year, y = lifeExp)) +
  geom_point()
#the very basic
```

```{r}
ggplot(data = gapminder, aes(x = year, y = lifeExp, colour = continent)) +
  geom_point()
#added color, well "coloUr"
```

this works so far
```{r}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, colour = continent)) + 
  geom_point()
#changed axes
```

```{r}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, colour = continent)) + 
  geom_point(aes(shape = continent))
#now with shapes!
```
HA I did it! This is part 1 on the HW

Plot 2
```{r}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, colour = continent)) + 
  geom_point(aes(shape = continent)) +
  scale_x_log10()
#shapes and log scale
```

```{r}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, colour = continent)) + 
  geom_point(aes(shape = continent)) +
  scale_x_log10() +
  stat_smooth(method = lm)
#added linear modelling lines, but went to all the sets.
```
closer, been working on it awhile, but can't get it to just one line for the 
 whole graph.....

Plot 3
```{r}
ggplot(data = gapminder, aes(x = gdpPercap, y = lifeExp, colour = continent)) + 
  geom_point(aes(shape = continent)) +
  scale_x_log10() +
  geom_smooth(aes(group = 1), size = 1, method = "lm", se = TRUE)
#grouped all the lines together
```


ggplot(data = gapminder, aes(x = lifeExp, y = , colour = continent)) + 
  geom_density(aes(shape = continent)) +
  facet_grid(year~lifeExp)
^ this one is broken
ggplot(data = gapminder, aes(x = lifeExp, y = , colour = continent, 
                             fill = continent)) + 
  geom_density(aes(shape = continent), alpha = 0.4)
^ this one is broken


```{r}
ggplot(data = gapminder, aes(x = lifeExp, y = , colour = continent, 
                             fill = continent)) + 
  geom_density(color = "black",alpha = 0.4)
#^ this one worked, but not faceted
```

```{r}
ggplot(data = gapminder, aes(x = lifeExp, y = , colour = continent, 
                             fill = continent)) + 
  geom_density(color = "black", alpha = 0.4) +
  facet_grid(year~.)
#^ this gets all horizontal facets. dammit. Can I swear in the comments? 
#is that kosher? Screw it. Sorry about your delicate sensibilities. You ass.
```

Plot 4
```{r}
ggplot(data = gapminder, aes(x = lifeExp, y = , colour = continent,
                             fill = continent)) + 
  geom_density(color = "black", alpha = 0.6) +
  facet_wrap(~year) 
```
Got it! With Jillian's help. stupid facet wrap.


ggplot(data = gapminder, aes(x = continent, y = lifeExp, colour = continent)) + 
  geom_boxplot() +
  facet_grid(year~continent)
 this is the box plot attempt
again, i think it's right, but wrong facet layout. the dots will squish 
together? need to figure out how to get all the boxes on the same graph?

Plot 5
```{r}
ggplot(data = gapminder, aes(x = continent, y = lifeExp, colour = continent)) + 
  geom_boxplot() +
  #this swaps density for box plots. 
  facet_wrap(~year) +
  theme(axis.text.x=element_text(angle=50, size=10, vjust=0.5))
#^ plot 5 and bonus! y axis labelling fixed.
```

Starting dplyer

select(gapminder)

tbl_df(gapminder)

group_by(gapminder, continent)

Plot 6
```{r}
ggplot(data = gapminder, aes(x = lifeExp, y = , colour = continent, 
                              fill = continent)) +
  geom_density(color = "black", alpha = 0.6)
```


Plot 7
```{r}
asian_data <- filter(gapminder, continent == "Asia")
#creating the data to pull from

ggplot(data = asian_data, aes(x = lifeExp, y = , colour = continent, 
                             fill = continent)) +
  scale_fill_manual(values = c("green")) +
  #I'm sure there's a way to not have the color messed with so much, 
  #but this works, even if ugly
  geom_density(color = "black", alpha = 0.6) + 
  geom_vline(aes(xintercept = mean(lifeExp))) +
  #adds the mean line
  theme(legend.position = 'none') + 
  #erases the legend
  ggtitle("Life expectancy in Asia")
#adds the title
#^ plot 7 got it!
```

8a Create a data.frame of the mean life expectancies for each continent
```{r}
mean_lifeExp <- 
  gapminder %>%
  group_by(continent) %>%
  summarize(avg = mean(lifeExp))


mean_lifeExp
```

mean_lifeExp <- 
  gapminder %>%
  group_by(continent) %>%
  summarize(avg = mean(lifeExp))

8b Plot the density plot of life expectancies for each continent 
 and draw a vertical line to mark the mean life expectancy for each continent.
```{r}
ggplot(data = gapminder, aes(x = lifeExp, y = , colour = continent, 
                             fill = continent)) +
  facet_wrap(~continent) + 
  geom_density(color = "black", alpha = 0.6) + 
  geom_vline(data = mean_lifeExp, aes(xintercept = avg))
#changes the data to the newly created data set and pulls the mean line
#from the avg created
```

Troubleshooting section
make sure to setwd("R_projects/")
then double check with list.files()
then run library(ggplot2)
can also check working directory with getwd()

original one:
mean_lifeExp <- mean(hw_gapminder$lifeExpe)

small_set <- hw_gapminder[c(1, 2, 3, 4, 1300:1304), ('country', 'continent', 'year')]

mean_gdp <- mean(hw_gapminder$gdpPercap)

max_country <- hw_gapminder$country[which(hw_gapminder$lifeExp = max(hw_gapminder$lifeExp))]

Fixed ones

```{r}
#extra "e" at the end
mean_lifeExp <- mean(hw_gapminder$lifeExp)
```

```{r}
#added "c" at front
small_set <- hw_gapminder[c(1, 2, 3, 4, 1300:1304), c('country', 'continent', 'year')]
```

```{r}
#add na.rm = TRUE
mean_gdp <- mean(hw_gapminder$gdpPercap, na.rm = TRUE)
```

```{r}
#missing extra =
max_country <- hw_gapminder$country[which(hw_gapminder$lifeExp == max(hw_gapminder$lifeExp))]
```

















